---
title: "Homework 1"
author: "Pradeeti Mainali"
date: "2024-09-18"
output: github_document
---

# Problem 1

Installed palmerpenguins in the console.

```{r}
# Loading the penguins data set:
data("penguins", package = "palmerpenguins")
```

### Description of penguins data set.

The data set contains `r nrow(penguins)` rows and `r ncol(penguins)` columns.

The penguin data set contains these variables: 
`r paste(names(penguins), collapse = ", ")`. 

The years of observation are 
`r min(penguins$year, na.rm = TRUE)`-`r max(penguins$year, na.rm = TRUE)`.
and the species include: 
`r paste(unique(penguins$species), collapse = ", ")`.

Islands include: 
`r paste(unique(penguins$island), collapse = ", ")`.


There are `r table(penguins$sex)["male"]` males 
and `r table(penguins$sex)["female"]` females

The mean bill length is `r mean(penguins$bill_length_mm, na.rm = TRUE)`,
the mean bill depth is `r mean(penguins$bill_depth_mm, na.rm = TRUE)`,
and the mean flipper length is `r mean(penguins$flipper_length_mm, na.rm = TRUE)`.

The body mass (in grams), ranging from `r min(penguins$body_mass_g, na.rm = TRUE)` 
to `r max(penguins$body_mass_g, na.rm = TRUE)` grams.

### Scatterplot of Flipper Length vs Bill Length

```{r penguin_plot, warning = FALSE}
library(ggplot2)

ggplot(data = penguins, 
  aes(x = bill_length_mm, y = flipper_length_mm, color = species)) +
  geom_point() +
  labs(title = "Scatterplot of Flipper Length vs Bill Length",
       x = "Bill Length (mm)",
       y = "Flipper Length (mm)")
ggsave("penguin_scatterplot.pdf", width = 8, height = 5)
```


