---
title: "Homework 1"
author: "Pradeeti Mainali"
date: "2024-09-18"
output: github_document
---

# Problem 1

Installed palmerpenguins in the console.

```{r}
# Loading the penguins data set:
data("penguins", package = "palmerpenguins")
```

### Description of penguins data set.

**The data set contains `r nrow(penguins)` rows and `r ncol(penguins)` columns.**

**The penguin data set contains these variables: **
**`r paste(names(penguins), collapse = ", ")`.**

The years of observation are 
`r min(penguins$year, na.rm = TRUE)`-`r max(penguins$year, na.rm = TRUE)`.
and the species include: 
`r paste(unique(penguins$species), collapse = ", ")`.

Islands include: 
`r paste(unique(penguins$island), collapse = ", ")`.


There are `r table(penguins$sex)["male"]` males 
and `r table(penguins$sex)["female"]` females

The mean bill length is `r mean(penguins$bill_length_mm, na.rm = TRUE)`,
the mean bill depth is `r mean(penguins$bill_depth_mm, na.rm = TRUE)`,
and **the mean flipper length is `r mean(penguins$flipper_length_mm, na.rm = TRUE)`**.

The body mass (in grams), ranging from `r min(penguins$body_mass_g, na.rm = TRUE)` 
to `r max(penguins$body_mass_g, na.rm = TRUE)` grams.

### Scatterplot of Flipper Length vs Bill Length

```{r penguin_plot, warning = FALSE}
library(ggplot2)

ggplot(data = penguins, 
  aes(x = bill_length_mm, y = flipper_length_mm, color = species)) +
  geom_point() +
  labs(title = "Scatterplot of Flipper Length vs Bill Length",
       x = "Bill Length (mm)",
       y = "Flipper Length (mm)")

ggsave("penguin_scatterplot.pdf", width = 8, height = 5)
```

# Problem 2

#### Loading in tidyverse:

```{r setup, message = FALSE}
library(tidyverse)
```

#### Creating the data set with random samples and vectors: 

```{r df_create}
set.seed(123)

p2_df=tibble(
  x = rnorm(10),
  vec_logical = x > 0,
  vec_char = sample(letters, 10, replace=TRUE),
  vec_factor = factor(
    sample(
      c("Level1", "Level2", "Level3"), 10, replace = TRUE))
)

print(p2_df)
```

#### Trying to explore the means of the variables:

```{r means_calc, warning = FALSE}

mean(pull(p2_df, var=x))

mean(pull(p2_df, var=vec_logical))

mean(pull(p2_df, var=vec_char))

mean(pull(p2_df, var=vec_factor))
```

When taking the mean of the variables in the data set,
we get 0.075 as the mean for x and 0.5 for vec_logical 
(could be because 50% are true and 50% are false). 
However, when trying to get the mean for vec_char and vec_factor, 
we get a warning message that the argument is not numeric or logical.

#### Trying the as numeric function

```{r numeric, warning = FALSE, results='hide'}

as.numeric(pull(p2_df, var=vec_logical))

as.numeric(pull(p2_df, var=vec_char))

as.numeric(pull(p2_df, var=vec_factor))
   
```

What happened when as.numeric was used:

* vec_logical: converted FALSE to 0 and TRUE to 1. R treated the logical value as binary data. 

* vec_char output NA's. Character were not / could not be converted to numbers since they were random characters, not binary, and did not represent any 'levels'

* vec_factor output numbers corresponding to the 'level' assigned to each observation. 

This helps understand what happens when we tried to take the mean of these variables. For the logical vector, the mean gave a proportion of TRUE values. For character and factor vector, it was unable to calculate the mean because of its original non-integer characteristic.

